---
title: "1.1 data wrangeling"
author: "TG"
date: "2022-10-30"
output: html_document
---


```{r}

library(tidyverse)
library(lubridate)
library(readxl)

```

# load data

```{r}

Aug_22_raw <- read_excel("data/Juvies August 2022.xlsx")
  
June_22 <- read_excel("data/Juvies June 2022.xlsx")
   
March_22 <- read_excel("data/Juvies March 2022.xlsx")

pre_data_raw<- read_csv("data/transects_data_for_analyses_2018-2020.csv")
```

# start orgenaize

fix date foramt to be uniform

```{r}
pre_data<-pre_data_raw

pre_data<-rename(.data = pre_data,"line"="...1")

pre_data<-pre_data %>% mutate(new_date = case_when(
  grepl("T00",pre_data$Trip.date) ==T ~ymd_hms(pre_data$Trip.date),
  grepl(" 0:00",pre_data$Trip.date) ==T ~dmy_hm(pre_data$Trip.date)))

pre_data$new_date<-ifelse(is.na(pre_data$new_date),as.character(dmy(pre_data$Trip.date)),
                          as.character(pre_data$new_date))

pre_data$new_date<-ymd(pre_data$new_date)

pre_data<-pre_data %>% relocate(new_date,.after = SiteNo) %>% select(-Trip.date) %>% rename("Trip_date" = "new_date",                                                                     "Site_name"="SiteNo",                                                                         "Sub_site"= "Site",                                                                           "Trans_type"="Letter",
       "Trans_surveyor_id"="ID",
       "Depth" = "Depth.m.",
       "Sp_status" = "c_t")

colnames(pre_data)


pre_data <- pre_data %>% mutate(survey_method = rep("Transects"),.after = month)

```

set Aug to the right format

- cover precentage
- object sizes
- transect id
- only transects

# Aug 22


```{r}

Aug_22<-Aug_22_raw

```

## delete unneccery columns

1. remove all duplicated columns with ".y"
2. calculate objects sizes (A_1,K_1...) and remove the excess columns
3. substrate %
4. column order and names


```{r}


# 1.
duplicated_col<- grep("\\.y$", colnames(Aug_22), ignore.case = T,value = T)

Aug_22<- Aug_22 %>% select(-duplicated_col)

Aug_22[,70:78] <- lapply(Aug_22[,70:78], as.character)
Aug_22[,70:78] <- lapply(Aug_22[,70:78], as.numeric)

Aug_22[,70:87][is.na(Aug_22[70:87])] <- 0

#2.
Aug_22$A1_size <- Aug_22$A1_height *Aug_22$A1_length *Aug_22$A1_width
Aug_22$A2_size <- Aug_22$A2_height *Aug_22$A2_length *Aug_22$A2_width
Aug_22$A3_size <- Aug_22$A3_height *Aug_22$A3_length *Aug_22$A3_width

Aug_22$K1_size = Aug_22$K1_height*Aug_22$K1_width*Aug_22$K1_length
Aug_22$K2_size = Aug_22$K2_height*Aug_22$K2_width*Aug_22$K2_length
Aug_22$K3_size = Aug_22$K3_height*Aug_22$K3_width*Aug_22$K3_length

Aug_22<- Aug_22 %>% select(-c(70:87))

#3.

per_cover_aug = Aug_22[,c(45:69)]

#there are 2 columns for s( S,s), unite:

per_cover_aug[per_cover_aug=="s"] = "S"  

#count how many times a letter appears in the selected columns:
per_cover_aug$S <- apply(per_cover_aug, 1, function(x) length(which(x=="S")))
per_cover_aug$SG <- apply(per_cover_aug, 1, function(x) length(which(x=="SG")))
per_cover_aug$C <- apply(per_cover_aug, 1, function(x) length(which(x=="C")))
per_cover_aug$G <- apply(per_cover_aug, 1, function(x) length(which(x=="G")))
per_cover_aug$B <- apply(per_cover_aug, 1, function(x) length(which(x=="B")))
per_cover_aug$R <- apply(per_cover_aug, 1, function(x) length(which(x=="R")))
per_cover_aug$A <- apply(per_cover_aug, 1, function(x) length(which(x=="A")))
per_cover_aug$SW = NA

per_cover_aug = per_cover_aug %>% mutate(S = (S/25) * 100,
                                 SG = (SG/25) * 100,
                                 C = (C/25) * 100,
                                 G = (G/25) * 100,
                                 B = (B/25) * 100,
                                 R = (R/25) * 100,
                                 A = (A/25) *100)


per_cover_aug = per_cover_aug %>% rowwise() %>% mutate(sum_check = sum(S,SG,C,G,B,R,A))

per_cover_aug = per_cover_aug %>% select(c(S,SG,C,G,B,R,A))

per_cover_aug$sum <- rowSums(per_cover_aug)

per_cover_aug <- per_cover_aug %>% mutate(across(c(1:7), ~ replace(.,sum == 0, NA)))

per_cover_aug$sum<-NULL  


#bind with data
Aug_22 = cbind.data.frame(Aug_22,per_cover_aug)

Aug_22<-Aug_22 %>% select(-c(45:69))


#4.

Aug_22<-Aug_22 %>% select(-Project,-Country,-SiteID,- Expedition)


 Aug_22<- Aug_22 %>% mutate(month = "august",
         year = "2022",
         Trans_No = paste(Transect,Letter))
 
Aug_22<- Aug_22 %>% mutate (observer = ifelse(Observer == "first", 
                           `First Observer.x`, 
                           `Second Observer.x`))

Aug_22<- Aug_22 %>% mutate(observer = case_when(
  Observer == "first" & Letter == "J" ~ `Juveniles First Observer.x`,
  Observer == "first" & Letter != "J" ~ `First Observer.x`,
  Observer == "second" & Letter == "J" ~ `Juveniles Second Observer.x`,
  Observer == "second" & Letter != "J" ~ `Second Observer.x`)) %>% relocate(observer,.after = Date)

 
#         observer = ifelse(Observer == "first", 
                           `First Observer.x`, 
                           `Second Observer.x`),
         trans_ID = paste(Site,Location.x,Transect,`First Observer.x`,`Second Observer.x`,month,year),
         ID = paste(Site,Location.x,observer,month,year)
         ) %>%
  #remove the observer column (not its name, but whether first/second) and create a column with the number of observers
  select(-c("Observer")) %>%
  group_by(trans_ID) %>%
  mutate(observer_num = ifelse(n_distinct(observer) > 1,
                               "two_observers",
                               "one_observer"),
         Depth = mean(c(`Depth Start`,`Depth End`),na.rm = T)) %>% # Calculate the mean depth between start and end points of the transect
  ungroup() %>%
  select(-c("Depth End","Depth Start"))



Aug_22<-Aug_22 %>% rename("survey_method" = "Method.x",
                          "Sub_site" = "Site",
                          "Site_name" = Location.x,
                          )

```




set March to the right format

```{r}

colnames(March_22)

March_22<-March_22 %>% select(-"Country",-"Project",-"Expedition",-"Juveniles Second Observer.y")
```

need to fix in predate:

* trip dates (v)
* NA coordinates

