---
title: "1.1 data wrangeling"
author: "TG"
date: "2022-10-30"
output: html_document
---


```{r}

library(tidyverse)
library(lubridate)
library(readxl)

```

# load data

```{r}

Aug_22_raw <- read_excel("data/Juvies August 2022.xlsx")
  
June_22_raw <- read_excel("data/Juvies June 2022.xlsx")
   
March_22_raw <- read_excel("data/Juvies March 2022.xlsx")

pre_data_raw<- read_csv("data/transects_data_for_analyses_2018-2020.csv")
```

# start orgenaize

fix date foramt to be uniform

```{r}
pre_data<-pre_data_raw

pre_data<-rename(.data = pre_data,"line"="...1")

pre_data<-pre_data %>% mutate(new_date = case_when(
  grepl("T00",pre_data$Trip.date) ==T ~ymd_hms(pre_data$Trip.date),
  grepl(" 0:00",pre_data$Trip.date) ==T ~dmy_hm(pre_data$Trip.date)))

pre_data$new_date<-ifelse(is.na(pre_data$new_date),as.character(dmy(pre_data$Trip.date)),
                          as.character(pre_data$new_date))

pre_data$new_date<-ymd(pre_data$new_date)

pre_data<-pre_data %>% relocate(new_date,.after = SiteNo) %>% select(-Trip.date) %>% rename("Trip_date" = "new_date",                                                                     "Site_name"="SiteNo",                                                                         "Sub_site"= "Site",                                                                           "Trans_type"="Letter",
       "Trans_surveyor_id"="ID",
       "Depth" = "Depth.m.",
       "Sp_status" = "c_t")

colnames(pre_data)


pre_data <- pre_data %>% mutate(survey_method = rep("Transects"),.after = month)

```

set Aug to the right format

- cover precentage
- object sizes
- transect id
- only transects

# Aug 22


```{r}

Aug_22<-Aug_22_raw

```

## delete unneccery columns

1. remove all duplicated columns with ".y"
2. calculate objects sizes (A_1,K_1...) and remove the excess columns
3. substrate %
4. column order and names


```{r}


# 1.
duplicated_col<- grep("\\.y$", colnames(Aug_22), ignore.case = T,value = T)

Aug_22<- Aug_22 %>% select(-duplicated_col)

Aug_22[,70:87] <- lapply(Aug_22[,70:87], as.character) #convert all of the A_hight... to numeric
Aug_22[,70:87] <- lapply(Aug_22[,70:87], as.numeric)

Aug_22[,70:87][is.na(Aug_22[70:87])] <- 0

#2.
Aug_22$A1_size <- Aug_22$A1_height *Aug_22$A1_length *Aug_22$A1_width
Aug_22$A2_size <- Aug_22$A2_height *Aug_22$A2_length *Aug_22$A2_width
Aug_22$A3_size <- Aug_22$A3_height *Aug_22$A3_length *Aug_22$A3_width

Aug_22$K1_size = Aug_22$K1_height*Aug_22$K1_width*Aug_22$K1_length
Aug_22$K2_size = Aug_22$K2_height*Aug_22$K2_width*Aug_22$K2_length
Aug_22$K3_size = Aug_22$K3_height*Aug_22$K3_width*Aug_22$K3_length

Aug_22<- Aug_22 %>% select(-c(70:87))

#3.

per_cover_aug = Aug_22[,c(45:69)]

#there are 2 columns for s( S,s), unite:

per_cover_aug[per_cover_aug=="s"] = "S"  

#count how many times a letter appears in the selected columns:
per_cover_aug$S <- apply(per_cover_aug, 1, function(x) length(which(x=="S")))
per_cover_aug$SG <- apply(per_cover_aug, 1, function(x) length(which(x=="SG")))
per_cover_aug$C <- apply(per_cover_aug, 1, function(x) length(which(x=="C")))
per_cover_aug$G <- apply(per_cover_aug, 1, function(x) length(which(x=="G")))
per_cover_aug$B <- apply(per_cover_aug, 1, function(x) length(which(x=="B")))
per_cover_aug$R <- apply(per_cover_aug, 1, function(x) length(which(x=="R")))
per_cover_aug$A <- apply(per_cover_aug, 1, function(x) length(which(x=="A")))
per_cover_aug$SW = NA

per_cover_aug = per_cover_aug %>% mutate(S = (S/25) * 100,
                                 SG = (SG/25) * 100,
                                 C = (C/25) * 100,
                                 G = (G/25) * 100,
                                 B = (B/25) * 100,
                                 R = (R/25) * 100,
                                 A = (A/25) *100)


per_cover_aug = per_cover_aug %>% rowwise() %>% mutate(sum_check = sum(S,SG,C,G,B,R,A))

per_cover_aug = per_cover_aug %>% select(c(S,SG,C,G,B,R,A))

per_cover_aug$sum <- rowSums(per_cover_aug)

per_cover_aug <- per_cover_aug %>% mutate(across(c(1:7), ~ replace(.,sum == 0, NA)))

per_cover_aug$sum<-NULL  


#bind with data
Aug_22 = cbind.data.frame(Aug_22,per_cover_aug)

Aug_22<-Aug_22 %>% select(-c(45:69))


#4.

Aug_22<-Aug_22 %>% select(-Project,-Country,-SiteID,- Expedition)


 Aug_22<- Aug_22 %>% mutate(month = "august",
         year = "2022",
         Trans_No = paste(Transect,Letter))
 
Aug_22<- Aug_22 %>% mutate (observer = ifelse(Observer == "first", 
                           `First Observer.x`, 
                           `Second Observer.x`))

Aug_22<- Aug_22 %>% mutate(observer = case_when(
  Observer == "first" & Letter == "J" ~ `Juveniles First Observer.x`,
  Observer == "first" & Letter != "J" ~ `First Observer.x`,
  Observer == "second" & Letter == "J" ~ `Juveniles Second Observer.x`,
  Observer == "second" & Letter != "J" ~ `Second Observer.x`)) %>% relocate(observer,.after = Date)



Aug_22<- Aug_22 %>% mutate(trans_ID = paste(Location.x,
                                            Site,
                                            Transect,
                                            Date,
                                            `First Observer.x`,
                                            `Second Observer.x`,
                                            `Juveniles First Observer.x`,
                                            `Juveniles Second Observer.x`
                                            ),
                           Trans_surveyor_id = paste(Location.x,Site,Transect,Date,observer))


# remove the observer column (not its name, but whether first/second) and create a column with the number of observers

Aug_22<-Aug_22 %>% select(-c(Observer,
                             `First Observer.x`,
                             `Second Observer.x`,
                             `Juveniles First Observer.x`,
                             `Juveniles Second Observer.x`))  


Aug_22<-Aug_22 %>% 
  group_by(trans_ID,Letter) %>%
  mutate(observer_num_adults = ifelse(n_distinct(observer) > 1 & Letter !="J" ,
                               "two_observers",
                               "one_observer"),
         observer_num_juv = ifelse(n_distinct(observer) > 1 & Letter =="J",
                               "two_observers",
                               "one_observer"),
         Depth = mean(c(as.numeric(`Depth Start`),as.numeric(`Depth End`)),na.rm = T)) %>% # Calculate the mean depth between start and end points of the transect
  ungroup() %>%
  select(-c("Depth End","Depth Start"))


Aug_22<-Aug_22 %>% rename("survey_method" = "Method.x",
                          "Sub_site" = "Site",
                          "Site_name" = "Location.x",
                          "Trans_type" = "Letter",
                          "lat"="Latitude",
                          "lon" = "Longitude")

Aug_22$line<-seq(1:nrow(Aug_22))


Aug_22$meta_to_deployment_id<-NULL
Aug_22$Comment <-rep(NA)

Aug_22<- Aug_22 %>% 
  mutate("Juv_depth_start" = case_when(
  `Juv Transect`=="A" ~ `A Depth Start`,
  `Juv Transect`=="B" ~ `B Depth Start`,
  `Juv Transect`=="C" ~ `C Depth Start`,
  `Juv Transect`=="D" ~ `D Depth Start`,
  `Juv Transect`=="E" ~ `E Depth Start`,
  `Juv Transect`=="F" ~ `F Depth Start`)) %>%
  mutate("Juv_depth_end" = case_when(
  `Juv Transect`=="A" ~ `A Depth End`,
  `Juv Transect`=="B" ~ `B Depth End`,
  `Juv Transect`=="C" ~ `C Depth End`,
  `Juv Transect`=="D" ~ `D Depth End`,
  `Juv Transect`=="E" ~ `E Depth End`,
  `Juv Transect`=="F" ~ `F Depth End`)) %>% 
  mutate("Juv_time_start" = case_when(
  `Juv Transect`=="A" ~ `A Time Start`,
  `Juv Transect`=="B" ~ `B Time Start`,
  `Juv Transect`=="C" ~ `C Time Start`,
  `Juv Transect`=="D" ~ `D Time Start`,
  `Juv Transect`=="E" ~ `E Time Start`,
  `Juv Transect`=="F" ~ `F Time Start`)) %>% 
  mutate("Juv_time_end" = case_when(
  `Juv Transect`=="A" ~ `A Time End`,
  `Juv Transect`=="B" ~ `B Time End`,
  `Juv Transect`=="C" ~ `C Time End`,
  `Juv Transect`=="D" ~ `D Time End`,
  `Juv Transect`=="E" ~ `E Time End`,
  `Juv Transect`=="F" ~ `F Time End`)) %>% select(-(8:31)) 


Aug_22<-Aug_22 %>% group_by(trans_ID,Trans_type) %>% 
  mutate(Depth_juv = mean(c(as.numeric(Juv_depth_start),as.numeric(Juv_depth_end)),na.rm = T)) %>% ungroup() %>% select(-Juv_depth_start,-Juv_depth_end)




```

for now ill remove the species specific columns from pre-data...

```{r}

pre_data<-pre_data %>% select(1:34)

pre_data<-pre_data %>% rename("Date" = "Trip_date",
                              "observer_num_adults"="observer_num",
                              "Amount"="Number_Ind")

pre_data$SW<-NULL
pre_data$Trans_No<-NULL
pre_data$Comment<-NULL


pre_data$sum <- rowSums(pre_data[21:27])


pre_data <- pre_data %>% mutate(across(c(21:27), ~ replace(.,sum == 0, NA)))

pre_data$sum<-NULL


```

```{r}

which(!colnames(pre_data) %in% colnames(Aug_22))

missing_col<-colnames(Aug_22[,(which(!colnames(Aug_22) %in% colnames(pre_data)))])

pre_data <-cbind(pre_data, setNames( lapply(missing_col, function(x) x=NA), missing_col))


which(!colnames(Aug_22) %in% colnames(pre_data))


Aug_22<-Aug_22 %>% select(colnames(pre_data))

Aug_22<- Aug_22 %>%
  relocate(observer_num_juv,.after = observer_num_adults) %>%
  relocate(Object,.after = Length) %>% 
  relocate(`Juv Transect`,.after = Transect) %>%
  relocate(`Time Start`,.after = Depth) %>%
  relocate(`Time End`,.before = Species) %>% 
  relocate(A2_size,.after = A1_size) %>%
  relocate(K3_size,.after = K2_size) %>% 
  relocate(A3_size,.before = K1_size) 

Aug_22<-Aug_22 %>% relocate(46:48,.after = 21)
Aug_22<-Aug_22 %>% relocate(46:47,.after = 36)

Aug_22<-Aug_22 %>% relocate(Notes,.after = Description)

pre_data<-pre_data %>% select(colnames(Aug_22))
```

bind Aug to the previous data

```{r}

Full_data<-rbind(Aug_22,pre_data)
Full_data$Comment<-NULL
Full_data$Trans_No<-NULL
```

ill first bind March and June  and then shape them together to fit the other data

```{r}

June_22<- June_22_raw
March_22<-March_22_raw

colnames(March_22[,(which(!colnames(March_22) %in% colnames(June_22)))])# check which columns not consistent  
colnames(June_22[,(which(!colnames(June_22) %in% colnames(March_22)))]) # check which columns not consistent  

June_22<-June_22 %>%
  select(-Method.y,-Location.y) %>% 
  rename("Location"="Location.x",
         "Method"="Method.x")

March_22$Method = rep("Transects")



March_22 <- March_22 %>% select(colnames(June_22)) # make column in the same order

June_march<- rbind(June_22,March_22)


```

ill repeat more or less the steps of Aug here...

```{r}

duplicated_col<- grep("\\.y$", colnames(June_march), ignore.case = T,value = T)

June_march<- June_march %>% select(-duplicated_col,-(92:109))

June_march[,57:74] <- lapply(June_march[,57:74], as.character)
June_march[,57:74] <- lapply(June_march[,57:74], as.numeric)

June_march[,57:74][is.na(June_march[57:74])] <- 0

#2.
June_march$A1_size <- June_march$A1_height *June_march$A1_length *June_march$A1_width
June_march$A2_size <- June_march$A2_height *June_march$A2_length *June_march$A2_width
June_march$A3_size <- June_march$A3_height *June_march$A3_length *June_march$A3_width

June_march$K1_size = June_march$K1_height*June_march$K1_width*June_march$K1_length
June_march$K2_size = June_march$K2_height*June_march$K2_width*June_march$K2_length
June_march$K3_size = June_march$K3_height*June_march$K3_width*June_march$K3_length


June_march$K1_size[June_march$K1_size==245000]<-0.245 # DATA WAS INPUR IN CM NOT IN M - FIX

June_march<- June_march %>% select(-c(57:74))

#3. calculate cover

per_cover_JM = June_march[,c(32:56)]

#there are 2 columns for s( S,s), unite:

per_cover_JM[per_cover_JM=="s"] = "S"  

#count how many times a letter appears in the selected columns:
per_cover_JM$S <- apply(per_cover_JM, 1, function(x) length(which(x=="S")))
per_cover_JM$SG <- apply(per_cover_JM, 1, function(x) length(which(x=="SG")))
per_cover_JM$C <- apply(per_cover_JM, 1, function(x) length(which(x=="C")))
per_cover_JM$G <- apply(per_cover_JM, 1, function(x) length(which(x=="G")))
per_cover_JM$B <- apply(per_cover_JM, 1, function(x) length(which(x=="B")))
per_cover_JM$R <- apply(per_cover_JM, 1, function(x) length(which(x=="R")))
per_cover_JM$A <- apply(per_cover_JM, 1, function(x) length(which(x=="A")))
per_cover_JM$SW = NA

per_cover_JM = per_cover_JM %>% mutate(S = (S/25) * 100,
                                 SG = (SG/25) * 100,
                                 C = (C/25) * 100,
                                 G = (G/25) * 100,
                                 B = (B/25) * 100,
                                 R = (R/25) * 100,
                                 A = (A/25) *100)


per_cover_JM = per_cover_JM %>% rowwise() %>% mutate(sum_check = sum(S,SG,C,G,B,R,A))

per_cover_JM = per_cover_JM %>% select(c(S,SG,C,G,B,R,A))

per_cover_JM$sum <- rowSums(per_cover_JM)

per_cover_JM <- per_cover_JM %>% mutate(across(c(1:7), ~ replace(.,sum == 0, NA)))

per_cover_JM$sum<-NULL  


#bind with data
June_march = cbind.data.frame(June_march,per_cover_JM)

June_march<-June_march %>% select(-c(32:56))


#4.

June_march<-June_march %>% select(-Project,-Country,-SiteID,- Expedition)

month(June_22_raw$Date[1])

June_march<- June_march %>% mutate(month = case_when(
  month(Date)==6 ~ "june",
  month(Date)==3 ~ "march",
  month(Date)==7 ~ "july"),
         year = "2022",
         Trans_No = paste(Transect,Letter))
 
June_march<- June_march %>% mutate (observer = ifelse(Observer == "first", 
                           `First Observer.x`, 
                           `Second Observer.x`))

June_march<- June_march %>% mutate(observer = case_when(
  Observer == "first" & Letter == "J" ~ `Juveniles First Observer.x`,
  Observer == "first" & Letter != "J" ~ `First Observer.x`,
  Observer == "second" & Letter == "J" ~ `Juveniles Second Observer.x`,
  Observer == "second" & Letter != "J" ~ `Second Observer.x`)) %>% relocate(observer,.after = Date)



June_march<- June_march %>% mutate(trans_ID = paste(Location,
                                            Site,
                                            Transect,
                                            Date,
                                            `First Observer.x`,
                                            `Second Observer.x`,
                                            `Juveniles First Observer.x`,
                                            `Juveniles Second Observer.x`
                                            ),
                           Trans_surveyor_id = paste(Location,Site,Transect,Date,observer))


# remove the observer column (not its name, but whether first/second) and create a column with the number of observers

June_march<-June_march %>% select(-c(Observer,
                             `First Observer.x`,
                             `Second Observer.x`,
                             `Juveniles First Observer.x`,
                             `Juveniles Second Observer.x`))  


June_march<-June_march %>% 
  group_by(trans_ID,Letter) %>%
  mutate(observer_num_adults = ifelse(n_distinct(observer) > 1 & Letter !="J" ,
                               "two_observers",
                               "one_observer"),
         observer_num_juv = ifelse(n_distinct(observer) > 1 & Letter =="J",
                               "two_observers",
                               "one_observer"),
         Depth = mean(c(as.numeric(`Depth Start`),as.numeric(`Depth End`)),na.rm = T)) %>% # Calculate the mean depth between start and end points of the transect
  ungroup() %>%
  select(-c("Depth End","Depth Start"))


June_march<-June_march %>% rename("survey_method" = "Method",
                          "Sub_site" = "Site",
                          "Site_name" = "Location",
                          "Trans_type" = "Letter",
                          "lat"="Latitude",
                          "lon" = "Longitude")

June_march$line<-seq(1:nrow(June_march))


June_march$meta_to_deployment_id<-NULL


June_march<- June_march %>% 
  mutate("Juv_depth_start" = case_when(
  `Juv Transect`=="A" ~ `A Depth Start`,
  `Juv Transect`=="B" ~ `B Depth Start`,
  `Juv Transect`=="C" ~ `C Depth Start`)) %>%
  mutate("Juv_depth_end" = case_when(
  `Juv Transect`=="A" ~ `A Depth End`,
  `Juv Transect`=="B" ~ `B Depth End`,
  `Juv Transect`=="C" ~ `C Depth End`)) %>% 
  mutate("Juv_time_start" = case_when(
  `Juv Transect`=="A" ~ `A Time Start`,
  `Juv Transect`=="B" ~ `B Time Start`,
  `Juv Transect`=="C" ~ `C Time Start`)) %>% 
  mutate("Juv_time_end" = case_when(
  `Juv Transect`=="A" ~ `A Time End`,
  `Juv Transect`=="B" ~ `B Time End`,
  `Juv Transect`=="C" ~ `C Time End`,
  )) %>% select(-(7:18)) 


June_march<-June_march %>% group_by(trans_ID,Trans_type) %>% 
  mutate(Depth_juv = mean(c(as.numeric(Juv_depth_start),as.numeric(Juv_depth_end)),na.rm = T)) %>% ungroup() %>% select(-Juv_depth_start,-Juv_depth_end)


```

```{r}
which(!colnames(Full_data) %in% colnames(June_march))
which(!colnames(June_march) %in% colnames(Full_data))

June_march$Comment<-NULL
June_march$Trans_No<-NULL
June_march$Habitat<-rep(NA)


June_march <- June_march %>% select(colnames(Full_data)) # make column in the same order

Full_data<- rbind(Full_data,June_march)

```

```{r}

Full_data$line<-NULL
Full_data<-Full_data %>% rename('juv_sub_transect'=`Juv Transect`)
Full_data<-Full_data %>% arrange(desc(Date),Site_name)

rm(list=setdiff(ls(), "Full_data"))
```


check foe missing coorinadtes,
in cryptic there is no cordinates yet so ill check for J or T


```{r}

Full_data$lat[Full_data$Site_name=="Igloo"] <-29.503679
Full_data$lon[Full_data$Site_name=="Igloo"] <-34.919032

Full_data$lat[Full_data$trans_ID =="Princess 4 A 2022-07-01 Tal Gavriel Avery Deveto Mai Lazarus Blank"] <-29.494216
Full_data$lon[Full_data$trans_ID =="Princess 4 A 2022-07-01 Tal Gavriel Avery Deveto Mai Lazarus Blank"] <-34.906752

coord_check<-Full_data %>% filter(Trans_type != "C",is.na(lat)) %>% distinct(Trans_surveyor_id,.keep_all = T)


```

34.919032

need to fix in predate:

* trip dates (v)
* NA coordinates

