---
title: "2.1 initial tests"
author: "TG"
date: "2022-11-14"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r}

library(tidyverse)
library(lubridate)
library(sf)
library(leaflet)
library(RColorBrewer)
library(stringr)
library(biogeo)
library(ggmap)
library(geosphere)
library(plotly)
library(TSP)
library(rgdal)
library(raster)

```

```{r}
load("Full_data.RData")



```


First lets create filtered data


```{r}

Adult_data<- Full_data %>% filter(Trans_type!= "J", survey_method == "Transects") %>% dplyr::select(-juv_sub_transect,-observer_num_juv,-survey_method)

```

data for test map - keep unique sample id only

```{r}

map_data<-Adult_data %>% distinct(trans_ID,.keep_all = T)

map_data$lat<-as.numeric(map_data$lat)
map_data$lon<-as.numeric(map_data$lon)



leaflet(map_data) %>% 
  addTiles() %>% 
  addAwesomeMarkers(lng = ~lon, lat = ~lat,label  = ~ trans_ID)
```


filter new surveyors data

```{r}
unique(Adult_data$observer)


Adult_data<-Adult_data %>% filter(observer != "Hava Wandel")
Adult_data<-Adult_data %>% filter(observer != "Avery Deveto") 
#for some reason their observation for transient species are not in the data!! need to be solved but ion the meantime...


```
for some reason their observation for transient species are not in the data!! need to be solved but ion the meantime ill just keep the other observer...

```{r}

Adult_data<-Adult_data %>% filter(!(observer == "Sarah Ohayon" & Site_name =="Migdalor"))
Adult_data<-Adult_data %>% filter(!(observer == "Sarah Ohayon" & Site_name =="Caves"))
Adult_data<-Adult_data %>% filter(!(observer == "Shevy Rothman" & Site_name =="Tour Yam"))
length(unique(Adult_data$trans_ID))

```

choose one observer randomly


```{r}
one_observer<-list()
set.seed(1)

for (i in unique(Adult_data$trans_ID)) {
  
  data <- Adult_data %>% filter(trans_ID == i)
  
  obs_amount <- length(unique(data$observer))
  obs_name <- unique(data$observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(observer == chosen_obs)
  one_observer[[i]]<-filter_data
  
  
}

one_observer <- bind_rows(one_observer)

length(unique(one_observer$trans_ID))

rm(data,filter_data)

```


# define my working data

```{r}
clean_adults<-one_observer
#save(clean_adults,file ="clean_adults.RData")
#write.csv(clean_adults,"clean_adults.csv")

```

# richness map


```{r}

clean_adults<-clean_adults %>% filter(Trans_type == "T") %>% filter(Confidence < 1)

length(unique(clean_adults$trans_ID))

species_richness <- clean_adults %>% group_by(trans_ID) %>% mutate(species_richness = n_distinct(Species))

species_richness$species_richness<-ifelse(species_richness$Species == "No Fish",0,species_richness$species_richness)

species_richness<-species_richness %>% distinct(trans_ID,.keep_all = T) %>% dplyr::select(1:14,species_richness)


```

Add all the transect where there were fish but the confidence >0 and there for are now "empty" - maybe i should consider changing sp richness to 1 in those cases...

```{r}
no_fish_transect <- unique(one_observer$trans_ID[which(!one_observer$trans_ID %in% clean_adults$trans_ID)])

no_fish <- one_observer %>% filter(trans_ID %in% no_fish_transect) %>% filter(Trans_type == "T") %>% distinct(trans_ID,.keep_all = T) %>% dplyr::select(1:14) %>% mutate(species_richness = 0)

noy_and_shahar<-one_observer %>% filter(trans_ID== "North Beach 2 B Shahar Chaikin Noy Shapira june 2020")

noy_and_shahar$Trans_type<-"T"

noy_and_shahar<-noy_and_shahar%>% dplyr::select(1:14) %>% mutate(species_richness = 0) # NO TRANSIENT DATA FOR SOME REASON (NO FISH THOGH)
```

unite the empty transects


```{r}

species_richness<-rbind(species_richness,no_fish)
species_richness<-rbind(species_richness,noy_and_shahar)

```


Static map

```{r}
backround_map = get_stamenmap(bbox = c(left =  34.900327,
                                       bottom = 29.490269,
                                       right = 34.983173,
                                       top = 29.554693),
                              zoom = 13,
                              maptype ="terrain-background")

backround_map<-ggmap(backround_map)

species_richness$lat<-as.numeric(species_richness$lat)
species_richness$lon<-as.numeric(species_richness$lon)

backround_map +
  geom_point(data = species_richness, aes(x = lon, y = lat, color = species_richness),alpha = 0.6,size = 2)+
   scale_color_gradient(low = "yellow", high = "red", na.value = NA)
```

Leaflet map

```{r}

color_pal<- colorNumeric(palette = "YlOrRd", domain = species_richness$species_richness,alpha = 0.8)

leaflet(species_richness) %>% 
  addTiles() %>% 
  addCircles(lng=~lon,lat=~lat,radius = 25,color = ~color_pal(species_richness),
    stroke = FALSE, fillOpacity = 0.8)

```
heatmap


```{r}
backround_map + stat_density2d(data=species_richness,
 aes(x=lon, y=lat, fill=..level..),alpha = 0.3, geom="polygon")+ 
  scale_fill_gradientn(colours=rev(brewer.pal(7, "Spectral")))
```


adu code

```{r}
#load libraries



#list work directories

wd_israel <- 'G:/My Drive/PHD/chapter 2  functional diversity/Eilat_FD/ISR_adm'


#load israeli map

israel <- readOGR('ISR_adm1', dsn = wd_israel)

#crop south

isr <- crop(israel, extent(34.26801, 35.90094, 29.49708, 29.57))

plot(isr, col = 'khaki1', bg = 'lightblue')

#load species richness

setwd(wd_sps_rich)

#species_richness <- readRDS('')

species_richness

#create spatialPointDataFrame

sps_rich_sp <- species_richness

sps_rich_sp <- sps_rich_sp[-which(is.na(sps_rich_sp$lon)),]

coordinates(sps_rich_sp) <- ~lon+lat

#plot points

plot(isr)
plot(sps_rich_sp, col = "magenta", cex = 0.5, pch = 19, add=T)


###################

#create empty raster

r_eilat <- raster(xmn = 34.86, xmx = 34.98, ymn = 29.48, ymx = 29.57, #extent
                  res = 0.0004, val = 0)


#create ID raster

ID_raster <- r_eilat
ID_raster[] <- c(1:length(ID_raster))

#create land raster
land_raster <- mask(ID_raster, isr) #crop ID raster by shp 
land_raster[!is.na(land_raster[])] <- 9999
land_raster[is.na(land_raster[])] <- 0
land_raster[which(land_raster[] == 9999)] <- NA

#identify observation site cells
pts_IDs <- extract(ID_raster, sps_rich_sp)
rich_ID <- as.data.frame(cbind(sps_rich_sp@data, raster_ID = pts_IDs))

#group

rich_ID<- rich_ID %>% group_by(raster_ID) %>% summarise("mean_rich" = mean(species_richness))

rich_ID2 <- as.data.frame(rich_ID)

#include values in raster

land_raster[rich_ID$raster_ID] <- rich_ID$mean_rich

plot(land_raster)



```

# Abundance map

